---
title: "Data Visualization, Lecture 4(Yen's notes)"
author: "Matt Wyczalkowski <m.wyczalkowski@wustl.edu>"
date: "Feb 8th, 2018"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

My favorite ggplot2 reference: http://www.cookbook-r.com/Graphs/

Load the ggplot2 library
```{r}
library(ggplot2)
```
# Read in BRFSS data
Get rid of all rows which have any missing values (NA)
```{r}
data.fn<-"dat/BRFSS.48K.csv"
BRFSS<-read.csv(data.fn)
BRFSS <- BRFSS[rowSums(is.na(BRFSS))==0,]
print(head(BRFSS))
```
# Bar Plot
## Basic plot of counts
Visualize employment status counts
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_bar(aes(x=employed))
p
```
Can't see the X-labels.

## Rotate X-label text
Jumping ahead a bit to themes
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_bar(aes(x=employed))
# Roate the X-axis text
p <- p + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) 
p
```
## Order columns
To change the order of the bars, change the order of factor levels by specifying the order explicitly.

### By an arbitrary order
```{r}
# Relevel the factors (instead of relevel function)
my.levels=c(
"Retired",
"Refused",
"A student",
"A homemaker",
"Self-employed",
"Unable to work",
"Employed for wages",
"Not asked or Missing",
"Out of work for 1 year or more",
"Out of work for less than 1 year"
)
BRFSS$employed <- factor(BRFSS$employed, levels = my.levels)
p <- ggplot(data=BRFSS)
p <- p + geom_bar(aes(x=employed))
p <- p + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
p
```


# Scatter Plot

## Basic scatter plot of weight vs. height
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=height, y=weight))
p
```
## Add color
We wish to indicate diabetes status with the color of the point.  
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=height, y=weight, color=diabetes_short_label))
p
```
## Use RColorBrewer Palette
Want to use RColorBrewer palette "set 1", with specified order of values

To see all RColorBrewer Palettes:
```{r}
library(RColorBrewer)
display.brewer.all()
```
```{r}
my.levels=c("NA",
"No",
"Yes",
"pre/borderline",
"Yes pregnancy",
"Refused",
"Unknown"
)
# Assign diabetes_shor_label 
BRFSS$diabetes_short_label <- factor(BRFSS$diabetes_short_label, levels = my.levels)
p <- ggplot(data=BRFSS)
# Use the RcolorBrewer palette="Set1"
p <- p + geom_point(aes(x=height, y=weight, color=diabetes_short_label)) + scale_color_brewer(palette="Set1")
p
```
## Change axis labels
Add units to axis labels.  Also, change legend name.
```{r}
my.levels=c("NA",
"No",
"Yes",
"pre/borderline",
"Yes pregnancy",
"Refused",
"Unknown"
)
BRFSS$diabetes_short_label <- factor(BRFSS$diabetes_short_label, levels = my.levels)
p <- ggplot(data=BRFSS)
#x, FULL. size, color
p <- p + geom_point(aes(x=height, y=weight, color=diabetes_short_label)) + scale_color_brewer(palette="Set1", name="Diabetes Status")
p <- p + xlab("Height [m]") + ylab("Weight [kg]")
p
```


# Violin Plot
What's the distribution of weight for different diabetes conditions?

## Plot weights as points
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=diabetes_short_label, y=weight))
p
```
This is hard to see because of overlap.  Let's jitter all points - add random displacement in X direction.  Also add color.
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=diabetes_short_label, y=weight, color=diabetes_short_label),size=0.1, position = position_jitter(w = 0.3, h = 0))
p <- p + scale_color_brewer(palette="Set1", name="Diabetes Status")
p
```
This is still hard to see because of overlapping points.  Reduce opacity (alpha) to make points more transparent
```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=diabetes_short_label, y=weight, color=diabetes_short_label), position = position_jitter(w = 0.3, h = 0), alpha=0.1)
p <- p + scale_color_brewer(palette="Set1", name="Diabetes Status")
p
```
Reduce alpha even more...
```{r}
# alpha: change the transparency
p <- ggplot(data=BRFSS)
p <- p + geom_point(aes(x=diabetes_short_label, y=weight, color=diabetes_short_label), position = position_jitter(w = 0.3, h = 0), alpha=0.01)
p <- p + scale_color_brewer(palette="Set1", name="Diabetes Status")
p
```
## Violin plot
Violin plots display the distribution (like a sideways histogram)

```{r}
#geom_violin
p <- ggplot(data=BRFSS)
p <- p + geom_violin(aes(x=diabetes_short_label, y=weight))
p
```
### Adjust colors and theme to make it pretty

```{r}
p <- ggplot(data=BRFSS)
#color=NA means transparent
p <- p + geom_violin(aes(x=diabetes_short_label, y=weight, fill=diabetes_short_label), color=NA) + guides(fill=FALSE)

p <- p + scale_fill_brewer(palette="Set1", name="Diabetes Status")
p <- p + theme_bw() + xlab("Diabetes Status") + ylab("Weight [kg]")
p <- p + theme(panel.grid.major.x = element_blank())
p
```

### Superpose additional geoms

```{r}
p <- ggplot(data=BRFSS)
p <- p + geom_violin(aes(x=diabetes_short_label, y=weight, fill=diabetes_short_label), color=NA) + guides(fill=FALSE)

p <- p + geom_point(aes(x=diabetes_short_label, y=weight, color=diabetes_short_label), position = position_jitter(w = 0.3, h = 0), alpha=0.01) + guides(color=FALSE)

p <- p + scale_fill_brewer(palette="Set1") + scale_color_brewer(palette="Set1", name="Diabetes Status") 
p <- p + theme_bw() + xlab("Diabetes Status") + ylab("Weight [kg]")
p <- p + theme(panel.grid.major.x = element_blank())
p
```

